\documentclass[../lecture-notes.tex]{subfiles}

\begin{document}
\begin{example}[revisited]
We have
\[
	\int_\Omega K(x, z) \phi(x) \sigma(x) \dd x = \gamma \phi(z)
\]
with $\sigma \equiv 1$, $K(x, z) = \min(x, z) - xz$ on $\Omega = [0, 1]$.
This gives 
\[
	\int_0^z x \phi(x) \dd x + \int_z^1 z \phi(x) \dd x - \int_0^1 xz \phi(x) \dd x = \gamma \phi(z).
\]
Now apply $L = - \frac{\diffd^2}{\diffd z^2}$ to this:
\begin{IEEEeqnarray*}{c"rCl}
	& \frac{\diffd^2}{\diffd z^2} \left\{ z \phi(z) - \int_1^z \phi(x) \dd x - z \phi(x) - \int_0^1 x \phi(x) \dd x \right\} &=& \gamma \phi''(z) \\
	\Longleftrightarrow & - \frac{1}{\gamma} \phi(z) &=& \phi''(z)
\end{IEEEeqnarray*}
\end{example}
\begin{remark}
From functional analysis one can derive, that for a Green's function which is positive semidefinite, Mercer's theorem applies, namely we obtain a representation by the eigenvalues and eigenfunctions.
One example would be Sturm-Liouville differential operators.

Note that not for all differential operators $L$ a Green's function will exist, e.g.\ if $L$ has a nontrivial null space of functions which also satisfy the boundary conditions.
\end{remark}
\begin{theorem} % Theorem 45
\label{thm:45}
Given a regularization operator $S$ with an expansion of $S^\dualst S$ into a discrete normalized eigendecomposition with $\gamma_i, \phi_i$, $i = 1, \ldots$ and a kernel with
\[
	K(x, z) \coloneqq \sum_{i, \gamma_i \neq 0} \frac{d_i}{\gamma_i} \phi_i(x) \phi(z),
\]
where $d_i \in \{ 0, 1\}$ for all $i$, and $\sum_{i=1}^\infty \frac{d_i}{\gamma_i}$ is convergent, then $K$ satisfies from \cref{thm:44}
\[
	\langle S K(x, \cdot), S K(z, \cdot) \rangle_D = K(x, z) = \langle K(x, \cdot), K(z, \cdot) \rangle_{\spH}.
\]
Moreover, the corresponding \ac{RKHS} is given by
\[
	\spn \{ \phi_i \mid d_i = 1, i \in \N \}.
\]
\end{theorem}
\begin{proof}
We use $f = K(z, \cdot)$ in \cref{thm:44}.
\begin{IEEEeqnarray*}{rCl}
\langle K(x, \cdot), S^\dualst S K(z, \cdot) \rangle &=& \left\langle \sum_{i} \frac{d_i}{\gamma_i} \phi_i(x) \phi_i(\cdot), S^\dualst S \left( \sum_{i} \frac{d_i}{\gamma_i} \phi_i(z) \phi_i(\cdot) \right) \right\rangle \\
&=& \sum_{i, j} \frac{d_i}{\gamma_i} \frac{d_j}{\gamma_j} \phi_i(x) \phi_j(z) \langle \phi_i(x), \underbrace{ S^\dualst S \phi_j(\cdot) }_{\gamma_j \phi_j} \rangle \\
&\overset{\text{orthonormal $\phi_i$}}{=}& \sum_{i} \frac{d_i}{\gamma_i^2} \gamma_i \phi_i(x) \phi_i(z) \\
&=& K(x, z)
\end{IEEEeqnarray*}
From the construction of the $K$ follows the statement about the span.
\end{proof}
Now going over to regularized empirical risks, we use the squared loss for $\ell$ and
\[
	s\left( \| f \|_{\spH} \right) = \frac{1}{2} \| f \|_{\spH}^2,
\]
i.e.
\[
	\min_{f \in \spH} \frac{1}{N} \sum_{i=1}^N \left( f(x_i) - y_i \right)^2 + \frac{\lambda}{2} \| f \|_{\spH}^2.
\]
We write
\[
	f = \sum_{j=1}^\infty \beta_j \phi_j
\]
after the theorem.
For simplicity consider positive definite operators, i.e.\ no eigenvalues are zero.
\begin{IEEEeqnarray*}{rCl}
\| f \|_{\spH}^2 = \langle f, f \rangle_{\spH} &=& \langle Sf, Sf \rangle_D \\
&=& \langle f, S^\dualst S f \rangle \\
&=& \Bigg\langle \sum_{j=1}^\infty \beta_j \phi_j, \underbrace{ S^\dualst S \sum_{j=1}^\infty \beta_j \phi_j }_{\sum_{i=1}^\infty \beta_j \gamma_j \phi_j} \Bigg\rangle \\
&=& \sum_{j=1}^\infty \beta_j^2 \gamma_j.
\end{IEEEeqnarray*}
Therefore,
\[
	\min_{f \in \spH} \riskRreg{\ell_2}(f) = \min_{f \in \spH} \frac{1}{N} \sum_{j=1}^N \left( f(x_i) - y_i \right)^2 + \frac{\lambda}{2} \sum_{j=1}^\infty \beta_j^2 \gamma_j.
\]
Differentiation after $\beta_j$ gives for $j= 1, \ldots, \infty$
\[
	\frac{\partial \riskRreg{\ell_2}}{\partial \beta_j} = \frac{1}{N} \sum_{i=1}^N (f(x_i) - y_i) \phi_j(x_i) + \beta_j \gamma_j \lambda.
\]
We define $\alpha_i \coloneqq y_i - f(x_i)$, i.e.\ pointwise unknowns.
Inserting this and solving for $\beta_j$ gives
\[
	\beta_j = \frac{1}{\lambda \gamma_j N} \sum_{i=1}^N \alpha_i \phi_j (x_i)
\]
and we obtain for $f$:
\begin{IEEEeqnarray*}{rCl}
f(x) = \sum_{j=1}^\infty \beta_j \phi_j &=& \frac{1}{\lambda}{N} \sum_{j=1}^\infty \frac{1}{\gamma_j} \sum_{i=1}^N \alpha_i \phi_j(x_i) \phi_j(x) \\
&=& \frac{1}{\lambda N} \sum_{i=1}^N \alpha_i K(x_i, x)
\end{IEEEeqnarray*}
with $K$ as in the theorem.
Inserting this representation into yields
\[
	\alpha_i = y_i - f(x_i) = y_i - \frac{1}{\lambda N} \underbrace{ \sum_{k=1}^N \alpha_k K(x_k, x_i) }_{(K\alpha)_i}.
\]
This gives in matrix notation:
\[
	\left( I + \frac{1}{\lambda N} K \right) \alpha = y
\]
or
\[
	\left( \tilde{\lambda} I + K \right) \underbrace{\frac{\alpha}{\tilde{\lambda}}}_{\tilde{\alpha}} = y
\]
with $\tilde{\lambda} = {\lambda}N$.
\begin{remark}
In a stochastic view, such an equation arises for Gaussian process regression. Here one models $y = f(x) + \varepsilon$, where $\varepsilon$ is an \ac{iid}\ Gaussian noise with variance $\sigma_N^2$. 
Then $K$ is viewed as a covariance function and the problem to solve is
\[
	\left( \sigma_N^2 I + K \right) \alpha = y.
\]
\end{remark}
How to solve $(\lambda I + K) \alpha = y$? Possibilities are
\begin{itemize}
\item LR decomposition
\item CG, maybe with preconditioners.
\begin{itemize}
\item When to stop the CG? Stopping has a regularization effect, called regularization by discretization\slash{}approximation.
\end{itemize}
\end{itemize}
\begin{remark}
Adding a scaled identity matrix improves the condition of the kernel matrix.
The better the condition, the ``worse'' are the results on the given data.
\end{remark}
We need to solve for several $\lambda$, therefore it is useful to compute once the eigendecomposition of $K$,
\[
	K = V \Gamma V^\tp,
\]
with an orthonormal matrix $V$ and $ \Gamma = \diag (x_i)_{i=1}^N$, where $x_i$ are the eigenvalues of $K$.
Then:
\begin{IEEEeqnarray*}{rCl}
	G \coloneqq (\lambda I + K)^{-1} &=& \left( \lambda I + V \Gamma V^\tp \right)^{-1} \\
	&=& V \left( \lambda I + \Gamma \right)^{-1} V^\tp.
\end{IEEEeqnarray*}
Having once performed the costly eigenvalue decomposition, we can solve for several $\lambda$'s using this equation:
\[
	\alpha^\lambda = V \left(\lambda I + \Gamma\right)^{-1} V^\tp y.
\]

In order to evaluate the performance of a predictive model, one usually uses cross-validation.
In the simplest setting, we split the available data $D$ into training data $D_T$ and the validation data $D_V$.
The empirical risk on $D_V$ is used to measure the predictive performance (or generalization) performance.

Usually, one uses $k$-fold \ac{CV}.
\begin{figure}[htpb]
\centering
\begin{tikzpicture}[scale=\textwidth/10cm]
	\foreach \x in {1,...,5}
		\draw ({2*\x}, 0) rectangle ({2*(\x+1)}, 1);
\end{tikzpicture}
\captionsetup{labelformat=empty}
\caption{Figure: Idea of $k$-fold \acl{CV}}
\end{figure}
We split $D$ into $k$ disjoint subsets. One is used as $D_V$, the other $k-1$ as $D_T$, solve and evaluate, repeat $k$ times.
The average over the $k$ empirical risks is the performance measure.
\end{document}