\documentclass[../skript.tex]{subfiles}

\begin{document}
\addtocounter{chapter}{-1}
\chapter{Optimal control of PDEs}
Consider optimization tasks in Banach spaces: \par
\begin{problem}
minimize $J(\xi)$ such that $\xi \in M$
\end{problem}
Here $J : X \to \R$ on a Banach space X, and $M \subseteq X$. Our motivation will be \textbf{optimal control problems}. \par
\textbf{Optimal control:} Physical processes modeled by differential or integral equations in function spaces involving a control variable to derive a certain cost function to a minimum.
\begin{example}
\begin{enumerate}
\item \textbf{Optimal stationary heating of a boundary:} \par
\textit{Target function:} $\Omega \subset \R^3$, $\Gamma$ boundary.
\begin{IEEEeqnarray*}{rCl}
J_\lambda : L^2(\Omega) \times L^2(\Gamma) &\to& \R, \lambda \geq 0 \\
(y, u) &\mapsto& \frac{1}{2} \int_\Omega (y(x) - y_\Omega(x))^2 \dx + \frac{\lambda}{2} \int_\Gamma |u(s)|^2 \ds \\
&& = \frac{1}{2} \| y - y_\Omega \|^2 + \frac{\lambda}{2} \| u \|^2 
\end{IEEEeqnarray*}
$y_\Omega$ is the target temperature, $y$ is called the state and $u$ the control. \par
\begin{problem}
minimize $J_\lambda(y, u)$, \ac{st}
\begin{IEEEeqnarray*}{r"l}
\begin{cases}
- \Delta y = 0 & \text{in } \Omega \\
\frac{\partial y}{\partial \nu} = \alpha(u - y) & \text{on } \Gamma
\end{cases} & \text{state equation} \\
u_a(x) \leq u(x) \leq u_b(x) & \text{restriction on control}
\end{IEEEeqnarray*}
\end{problem}
In abstract terms: 
\begin{problem}
minimize $J_\lambda(y, u)$, \ac{st}
\begin{IEEEeqnarray*}{rCl}
Ay &=& Bu \\
u_a &\leq& u \leq u_b
\end{IEEEeqnarray*}
\end{problem}
\begin{itemize}
\item $J_\lambda$ is strictly convex (even quadratic)
\item constraints constitute a convex set $M$
\end{itemize}
Thus, we have a convex optimization in a Hilbert space.
By restricting to a smaller function space (Sobolev space), one may achieve that $y$ depends uniquely on $u$:
\begin{equation}
y = A^{-1} B u
\end{equation}
This given a \textbf{reduced problem}: \par
\begin{problem}
minimize $f_\lambda(u) = J(A^{-1} B u)$ \par
\ac{st} $u_a \leq u \leq u_b$
\end{problem}
Minimizer $\bar{u}$ is called optimal control. \par
\textbf{Questions:} Existence, uniqueness, necessary conditions, optimization methods.
\item \textbf{Instationary boundary heating} \par
Achieve specific temperature at fixed time $T$ (``hot potato'').
Now $y = y(x, t)$, $u = u(x, t)$.
\begin{problem}
minimize $J_\lambda(y, u)$
\begin{IEEEeqnarray*}{rCl}
J_\lambda(y, u) &=& \frac{1}{2} \int_\Omega (y(x, T) - y_\Omega(x))^2 \dx + \frac{\lambda}{2} \int_0^T \int_\Gamma |u(x, t)|^2 \ds
\end{IEEEeqnarray*}
\ac{st}
\begin{IEEEeqnarray*}{l}
\left\{\begin{IEEEeqnarraybox}[][c]{rCl"l}
y_t - \Delta y &=& 0 & \text{in } Q = \Omega \times (0, T) \\
\frac{\partial y}{\partial \nu} &=& \alpha(u - y) & \text{in } \Sigma = \Gamma \times (0, T) \\
y(x, 0) &=& y_0(x) & \text{in } \Omega
\end{IEEEeqnarraybox}\right. \\
u_a(x, t) \leq u(x, t) \leq u_b(x, t)
\end{IEEEeqnarray*}
\end{problem}
Again we have a convex problem.
\item \textbf{Optimal stationary heating inside}
Heat source $E \subseteq \Omega$
\begin{IEEEeqnarray*}{l}
\begin{cases}
- \Delta y = \chi_E \cdot u & \text{in } \Omega \\
y = 0 & \text{on } \Gamma
\end{cases}
\end{IEEEeqnarray*}
\end{enumerate}
\end{example}
\begin{example}[Non-convex examples]
Non-convexity arises in optimal control by
\begin{itemize}
\item non-convex target functional $J$
\item non-convex state equation
\item non-convex control restrictions
\end{itemize}
\begin{enumerate}
\setcounter{enumi}{3}
\item \textbf{Heating with radiation:} \par
Like in the first example, let
\begin{IEEEeqnarray*}{l}
\begin{cases}
- \Delta y = 0 & \text{in } \Omega \\
\frac{\partial y}{\partial \nu} = \alpha(u^4 - y^4) & \text{on } \Gamma
\end{cases}
\end{IEEEeqnarray*}
Semilinear elliptic PDE.
\item \textbf{Supra-conduction:} \par
Simplified Ginzburg-Landan equation
\begin{IEEEeqnarray*}{l}
\begin{cases}
- \Delta y - y + y^3 = u & \text{in } \Omega \\
y = 0 & \text{on } \Gamma
\end{cases}
\end{IEEEeqnarray*}
\end{enumerate}
\end{example}
\paragraph{Main concepts (in finite dimension)}
Let $f : \R^n \to \R$ be $C^1$, strictly convex and \textbf{coercive}, that is $f(\xi) \to + \infty$ if $\| \xi \| \to + \infty$.
Assume $M \subset \R^n$ is closed and convex. Consider
\begin{problem}
minimize $f(\xi)$ \ac{st} $\xi \in M$.
\end{problem}
\begin{theorem}
This problem has a unique solution $\bar{\xi}$.
\end{theorem}
\begin{proof}
Pick $\xi_0 \in \R^n$, then
\[
\mathcal{L}_0 = \{ \xi \in \R^n : f(\xi) \leq f(\xi_0) \}
\]
is compact. Therefore $\mathcal{L}_0 \cap M$ is compact and a solution $\bar{\xi}$ exists.
Let $\bar{\xi}, \bar{\bar{\xi}} \in M, \bar\xi \neq \bar{\bar{\xi}}$ be two solutions. Then the line segment $\ell$ connecting $\bar\xi$ and $\bar{\bar{\xi}}$ belongs to $M$. In particular,
\[
\hat{\xi} = \bar\xi + \frac{1}{2} \left(\bar{\bar{\xi}} - \bar\xi\right) = \frac{1}{2} \bar{\bar{\xi}} + \left(1 - \frac{1}{2} \right) \bar\xi
\]
but
\[
f\left(\hat{\xi}\right) < \frac{1}{2} f \left(\bar{\bar{\xi}}\right) + \left(1 - \frac{1}{2}\right) f\left(\bar{\xi}\right) = f\left(\bar{\xi}\right).
\]
Contradiction to strict convexity.
\end{proof}
\textbf{Note:}
\begin{itemize}
\item Coercivity of $f$ + closeness of $M$ $\to$ existence
\item strict convexity of $f$ + convexity of $M$ $\to$ uniqueness
\end{itemize}
\paragraph*{Optimality conditions:}
\textbf{Tangent cone} of convex $M$:
\[
T_{\bar{\xi}}(M) = \{ \alpha(\xi - \bar{\xi}) \mid \alpha \geq 0, \xi \in M \}
\]
\textbf{Normal cone}:
\begin{IEEEeqnarray*}{rCl}
N_{\bar{\xi}} &=& (T_{\bar{\xi}}(M))^\circ = (\conv(T_{\bar{\xi}}(M)))^\circ \\
&=& \{ v : \langle v, h \rangle \leq 0 \quad \forall h \in T_{\bar{\xi}} (M) \}
\end{IEEEeqnarray*}
\textbf{Optimality condition:} 
\begin{IEEEeqnarray*}{r"rCl"l'l}
& - \nabla f(\bar{\xi}) &\in& N_{\bar{\xi}}(M) \\
\Leftrightarrow & \langle \nabla f(\bar{\xi}), h \rangle &\geq& 0 & \forall h \in T_{\bar{\xi}}(M) & \textbf{(Variational inequality)} \\
\Leftrightarrow & \langle \nabla f(\bar{\xi}), \xi - \bar{\xi} \rangle &\geq& 0 & \forall \xi \in M
\end{IEEEeqnarray*}
\paragraph*{Prototypical control theory problems}
\begin{problem}
minimize $J(y, u)$, $y \in \R^n$, $u \in \R^m$, \ac{st}
\begin{IEEEeqnarray*}{c}
Ay = Bu, \; A \text{ is invertible} \\
\underbrace{u_a \leq u \leq u_b}_{\text{``box constraint''}} \Leftrightarrow u \in U_{ad}
\end{IEEEeqnarray*}
\end{problem}
Then:
\[
M = \left\{ \begin{pmatrix}
y \\ u
\end{pmatrix} = \begin{pmatrix}
A^{-1} B u \\ u 
\end{pmatrix} : u \in U_{ad} \right\}
\]
is convex. \par
\textbf{Variational inequality:} if $(\bar{y}, \bar{u})$ solves the problem then
\begin{IEEEeqnarray*}{r"rCl"l}
& \langle \nabla_y J(\bar{y}, \bar{u}), A^{-1} B (u - \bar{u}) \rangle + \langle \nabla_u J(\bar{y}, \bar{u}), u - \bar{u} \rangle &\geq& 0 & \forall u \in U_{ad} \\
\Leftrightarrow & \langle B^\tp A^{-\tp} \nabla_y J(\bar{y}, \bar{u}) + \nabla_u J(\bar{y}, \bar{u}), u - \bar{u} \rangle &\geq& 0 & \forall u \in U_{ad} \IEEEyesnumber \label{eq:c0e1}
\end{IEEEeqnarray*}
\textbf{Lagrange function for equality constraint:}
\[
L(y, u, p) = J(y, u) - \langle Ay - Bu, p \rangle, p \in \R^n.
\]
Then,
\[
\nabla_y L(y, u, p) = 0
\]
is equivalent to
\[
A^\tp p = \nabla_y J(y, u),
\]
(because $\langle \nabla_y J(\bar{y}, \bar{u}), h \rangle = \langle Ah, p \rangle$) which is called \textbf{adjoint equation} and
\[
p = A^{-\tp} \nabla_y J(y, u)
\]
is called \textbf{adjoint state}. Therefore \eqref{eq:c0e1} reads
\begin{IEEEeqnarray*}{rCl"l}
\nabla_y L(\bar{y}, \bar{u}, \bar{p}) &=& 0 \\
\langle \nabla_u L(\bar{y}, \bar{u}, \bar{p}), u - \bar{u}) &\geq& 0 & \forall u \in U_{ad}
\end{IEEEeqnarray*}
\textbf{Note:} $\bar{p}$ is a Lagrange multiplier. \par
\textbf{Lagrange function for inequality constraints:}
\[
u_a - u \leq 0, u - u_b \leq 0.
\]
Let
\[
L(y, u, p, \mu_a, \mu_b) = J(y, u) - \langle Ay - Bu, p \rangle + \langle u_a - u, \mu_a \rangle + \langle u - u_b, \mu_b \rangle
\]
\begin{theorem}
If $(\bar{y}, \bar{u})$ is a solution, then there exists $\bar{p} \in \R^n$, $\bar{\mu}_a \in \R^m$, $\bar{\mu}_b \in \R^m$, \ac{st}
\[
\begin{cases}
\begin{IEEEeqnarraybox}{rCl}
\nabla_y L(\bar{y}, \bar{u}, \bar{p}, \bar{\mu}_a, \bar{\mu}_b) &=& 0 \\
\nabla_u L(\bar{y}, \bar{u}, \bar{p}, \bar{\mu}_a, \bar{\mu}_b) &=& 0 \\
\bar{\mu}_a \geq 0, \bar{\mu}_b &\geq& 0 \\
\langle u_a - \bar{u}, \bar{\mu}_a \rangle = 0, \langle \bar{u} - u_b, \bar{\mu}_b \rangle &=& 0
\end{IEEEeqnarraybox}
\end{cases}
\]
These conditions are called \textbf{KKT conditions}.
\end{theorem}
\begin{proof}
Choose
\[
\bar{\mu}_a \coloneqq (B^\tp \bar{p} + \nabla_u J(\bar{y}, \bar{u}))_+ = (\nabla_u L(\bar{y}, \bar{u}, \bar{p}))_+
\]
and
\[
\bar{\mu}_b \coloneqq (-[\nabla_u L(\bar{y}, \bar{u}, \bar{p})])_+.
\]
\end{proof}
\paragraph{Plan:}
\begin{enumerate}
\item Optimality in Banach spaces
\begin{itemize}
\item weak compactness, convexity
\item dual spaces, Lagrange multipliers
\end{itemize}
\item Application to PDE control problems
\begin{itemize}
\item theory for PDE
\item Sobolev spaces
\end{itemize}
\item Optimization methods
\end{enumerate}
\end{document}