\documentclass[../skript.tex]{subfiles}

\begin{document}
\begin{definition}
An \emph{approximation algorithm} $A$ for some optimization problem $P$ is a polynomial time algorithm for $P$. We say $A$ has \emph{approximation ratio} $k$ $(k \in \R)$ if
\begin{IEEEeqnarray*}{rCl'l}
	A(I) &\leq& k \OPT(I), & P \text{ is a minimization problem} \\
	A(I) &\geq& k \OPT(I), & P \text{ is a maximization problem}
\end{IEEEeqnarray*}
For minimization problems $k \geq 1$ and for maximization problems $k \leq 1$.
\end{definition}
\begin{problem}[$k$-Sat]
\begin{tabular}{@{}ll}
\textit{Input:} & Instance of \textsc{Sat} such that each clause contains at most $k$ literals. \\
\textit{Question:} & Is there a truth assignment for $F$?
\end{tabular}
\end{problem}
\begin{theorem} % Thm 4
\label{thm:4}
$k$-\textsc{Sat} is \NP-complete for $k \geq 3$.
\end{theorem}
\begin{proof}
$k$-\textsc{Sat} $\in \NP$ is clear.
It suffices to show that $\textsc{3Sat}$ is \NP-complete. Reduce \textsc{Sat} to \textsc{3Sat}.
Let $F$ be an instance of $\textsc{Sat}$, $F = C_1 \wedge \ldots \wedge C_k$.
Choose a clause $C_i = a_1 \vee \ldots \vee a_l$ with $l \geq 4$.
Consider that
\[
(a_1 \vee \ldots \vee a_l) \; \Leftrightarrow \; (a_1 \vee a_2 \vee x) \wedge (a_3 \vee a_4 \ldots \vee a_l \vee \bar{x})
\]
By iterating this for every clause, we obtain a \textsc{3Sat} instance equivalent to the original instance.
\end{proof}
\begin{theorem} % Thm 5
\label{thm:5}
\textnormal{2\textsc{Sat}} is polynomially solvable
\end{theorem}
\begin{proof}
Let $F = C_1 \wedge \ldots \wedge C_k$ be an instance of 2\textsc{Sat}. Construct a digraph $G = (V, E)$ with vertices corresponding to literals of $F$.
For a clause $(a \vee b)$ insert edges $(\bar{a}, b)$ and $(\bar{b}, a)$.
\underline{Claim:} $F$ has a truth assignment \ac{iff} in $G$ there is no $x$ for which $\bar{x}$ is in the same strongly connected component.
\begin{itemize}
\item[$(\Rightarrow)$:] Existence of a path from $x$ to $\bar{x}$ implies $x$ has to be \textit{false} for any valid truth assignment. Conversely, $x$ would have to be \textit{true} as well, which is not possible.
\item[$(\Leftarrow)$:] Contract all strongly connected component and obtain $\tilde{G}$. Choose a vertex $\tilde{v} \in V(\tilde{G})$ without any incoming edges. Set all variables in the set $H$ corresponding to $\tilde{v}$ in $G$ to \textit{false}. Similarly, for the graph of the negated literals of $H$, $\bar{H}$, set the literals to true. Now, remove $H$ and $\bar{H}$ from $\tilde{G}$.
\end{itemize}
\end{proof}
\begin{problem}[Directed Hamiltonian Cycle]
\begin{tabular}{@{}ll}
\textit{Input:} & $G = (V, E)$ directed graph. \\
\textit{Question:} & Does $G$ contain a cycle visiting all vertices?
\end{tabular}
\end{problem}
\begin{theorem} % Thm 6
\label{thm:6}
\textsc{Directed Hamiltonian Cycle} is \NP-complete
\end{theorem}
\begin{proof}
The problem is obviously in \NP.
Reduce to \textsc{3Sat} to \textsc{Directed Hamiltonian Cycle}. Let $F = C_1 \wedge \ldots \wedge C_k$ with variables $x_1, \ldots, x_n$.
For each variable, introduce a path of length $3 \cdot k + 2$ directed in both directions.
Add a vertex for each clause. Add vertices $s, t$ and connect
\begin{itemize}
\item $s$ to both endpoints of the first path
\item every path endpoints to the endpoints of the next path
\item each endpoint of the last path to $t$
\item $t$ to $s$
\end{itemize}
Now we distinguish:
\begin{itemize}
\item If $x_i$ is in clause $C_j$. Pick the $j$-th edge $e = (u, v)$ on the path of $x$ and connect $u$ to $C_j$ and $C_j$ to $v$.
\item If $\bar{x_i}$ is in $C_j$. Perform the same procedure in the other direction.
\end{itemize}
\underline{Claim:} $F$ is satisfiable \ac{iff} $G$ has a directed Hamiltonian cycle.
\begin{proof}
\begin{itemize}
\item[$(\Rightarrow)$] Choose a truth assignment. Traverse $P_i$ from
\[
\begin{cases}
\text{left to right}, & \text{if } x_i \text{ is \textit{true}}. \\
\text{right to left}, & \text{if } x_i \text{ is \textit{false}}.
\end{cases}
\]
Choose one true literal per clause $C_j$. Add vertex $C_j$ in the path corresponding to that literal.
\item[$(\Leftarrow)$] Let $G$ have a Hamiltonian cycle $H$. \\
\underline{Claim:} If $H$ reaches $C_j$ from path $P_i$, then $H$ continues along $P_i$.
\begin{proof}
Assume the path visited $C_j$ from $\beta_j$ and proceeds to $\gamma_j$. Let $\alpha_j$ be the predecessor of $\beta_j$.
\begin{itemize}
\item $\alpha_j$ is only connected with its neighbors on $P_i$, so $H$ has to come from the left on $P_i$.
\item $\gamma_j$ can only be visited from $C_j$.
\end{itemize}
\end{proof}
\end{itemize}
\end{proof}
\end{proof}
\begin{problem}[Hamiltonian Cycle]
\begin{tabular}{@{}ll}
\textit{Input:} & $G = (V, E)$ undirected graph. \\
\textit{Question:} & Is there a cycle in $G$ containing all vertices?
\end{tabular}
\end{problem}
\begin{theorem} % Thm 7
\label{thm:7}
\textsc{Hamiltonian Cycle} is \NP-complete.
\end{theorem}
\begin{proof}
\textsc{Hamiltonian Cycle} being in \NP{} is obvious.
Reduce \textsc{Directed Hamiltonian Cycle} to \textsc{Hamiltonian Cycle}.
Start with an empty graph $G'$. For each vertex $v \in V(G)$ three vertices $v'$, $v''$, $v'''$ are added. Connect $v'$ to $v''$ and $v''$ to $v'''$. Incoming edges of $v$ are connected to $v'$, outgoing edges are connected to $v'''$.
Now $G$ has a \textsc{Directed Hamiltonian Cycle} \ac{iff} $G'$ has a \textsc{Hamiltonian Cycle}. \\
\underline{Observation:} On a Hamiltonian Cycle for $G'$, the vertices $v'$, $v''$ and $v'''$ are traversed consecutively.
\end{proof}
\begin{problem}[Vertex Cover]
\begin{tabular}{@{}ll}
\textit{Input:} & $G = (V, E)$ undirected graph, $k \in \N$. \\
\textit{Question:} & Is there a set $S \subseteq V$ with $|S| \leq k$, \ac{st} each edge has \\
& an endpoint in $S$?
\end{tabular}
\end{problem}
\begin{theorem} % Thm 8
\label{thm:8}
\textsc{Vertex Cover} is \NP-complete.
\end{theorem}
\begin{proof}
\textsc{Vertex Cover} being in \NP{} is obvious.
Reduce \textsc{3Sat} to \textsc{Vertex Cover}.
Let $F = C_1 \wedge \ldots \wedge C_k$. Construct a graph $G = (V, E)$
\Ac{wlog} each clause contains exactly 3 literals.
For each clause $C_i$ construct a clique with vertices corresponding to its 3 literals.
Add an edge between each $x$, $\bar{x}$.

\underline{Claim:} $F$ has a truth assignment \ac{iff} $G$ contains a \textsc{Vertex Cover} with at most $2k$ vertices.
\begin{itemize}
\item[$(\Rightarrow)$] Choose one true literal in each clause. Set $S$ as the set of all other vertices.
\item[$(\Leftarrow)$] Let $S$ be a vertex cover, $S \leq 2k$. Each clique needs to contain 2 vertices from $S$. Thus, there are exactly 2 vertices per clique in $S$.
Set the literals \emph{not} corresponding to vertices in $S$ to \textit{true}.
\end{itemize}
\end{proof}
\begin{problem}[Stable Set]
\begin{tabular}{@{}ll}
\textit{Input:} & $G = (V, E)$ undirected graph, $k \in \N$. \\
\textit{Question:} & Is there a set $S \subseteq V$ with $|S| \geq k$, \\
& \ac{st} there are no edges connecting vertices in $S$?
\end{tabular}
\end{problem}
\begin{theorem} % Thm 9
\label{thm:9}
\textsc{Stable Set} is \NP-complete.
\end{theorem}
\begin{proof}
It is obvious that \textsc{Stable Set} is in \NP.
Reduce \textsc{Vertex Cover} to \textsc{Stable Set}.
If $G = (V, E)$ is an instance of \textsc{Vertex Cover}, then we can find a stable set $S$ in this graph: $V \setminus S$ is a stable set.
Each vertex cover is the complement of a stable set.
\end{proof}
\end{document}